On the eletrodynamic of moving bodies

### 2019-06-05 Article_AS.Rmd
Intro version1

---
title: "On the eletrodynamic of moving bodies"
author: "André Dogninema Soro, Jean-Romain Roussel, Alexis Achim"
editor_options:
  chunk_output_type: inline
output:
  pdf_document: default
  word_document: default
bibliography: biblio.bibtex
---

@@

Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of Canadian forest are harvested each year and an additional loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain resource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better structural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \

@@

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood quality has a biologic and an economic component. Biologically, the resistance to mortality by vigour is important to ensure to harvest leaving trees. Economically the value of the wood for end-use is important for industries. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. More specifically the homogeneity of the density within the stem has an high importance. I has been demonstrated that trees with a more uniform wood density within the stem are also the more resistant to mortality [@Hietz2013; @osazuwa2014] because a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. In fact, wood density uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bio energy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For example, a strong inter-annual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance [@x,@y]. It is therefore important from both a biological and an end-use points of view to have a wood with a uniform density within the stem. \

@@

### L'importance de la prise en compte du climat et de l'héritabilité des caractères dans les les programmes de selection génétique
<!-- Ca part en couille il n'y a plus de logique. Quel message veux tu faire passer ? Tu te tires une balle dans le pied en disant l'inverse de que tu as dit avant ! --> 
The time-frame of improved forest management being long [@Zobel1972] it is important to anticipate future changes in environmental conditions in these programs [@Gray2016]. This will allow the establishment of plantations capable of adapting to the water stress and possible significant rainfall related to climate change because climate predictions predict more variations in the coming years [@IPCC2007; @pachauri2014]. In fact, with the increasing temperature, the trees are expected to growth faster and thus being of lower quality because of lower density [@corriveau1987; @DeBell1994; @Lindstrom1996; @Dutilleul1998; @Bouriaud2005] hence, the importance of integrating these factors into genetic selection programs. However, to select trees that are prone to produce wood of homogeneous density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accurately, with climate change. Furthermore, heritability of density has been proven in several studies, but there is no evidence to suggest that density variability is hereditary. The heritability of this trait could qualify it for the selection criteria for future breeding programs for producing better quality wood.\

### Le but de notre étude et ce que nous avons fait. 
In this study we quantified and compared heterogeneity of wood density from trees of known parents in order to figure out if some individuals produce more heterogeneous wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the heterogeneity by comparing two growth sites we tried to characterize the environmental contribution to homogeneity. We ultimately aim to select and reproduce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogeneous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

@@

## Sampling
The dataset used in this study came from genetic selection for a series of tests according to provenances [@Beaulieu1996]. The improvement program for this region was led by Quebec’s Ministry of Forests, Wildlife and Parcs, Canada. Experimental plantings were initiated from white spruce seedlings grown in nurseries and planted at two years of age at a regular spacing between trees of 2 m in all directions. This series of tests was established in 1999 by the Canadian Forest Service and is one of the three essential parts of the white spruce breeding program in Quebec. Two sites were selected for sampling: one near Saint-Casimir, Quebec (46.6994, -72.1119), in the maple–basswood ecological zone (warm weather) , and the other one near Cabano (Asselin), Quebec (47.8808, -68.4436), in the balsam fir-yellow birch ecological zone (cool weather) (fig 1). Between 9 and 14 trees per site were sampled from each of the 93 full-sib families, for a total sample of 2365 trees. Tree height and diameter at breast height (DBH, 1.3 m above ground) were recorded in 2015 at the age of 15 years, and a wood increment core was extracted from the south facing side of each tree. Cores were stored in a freezer, conditioned to 7% moisture and cut to 1.68 mm thickness prior to X-ray densitometry analyses (Quintek Measurement Systems, TN). Wood density was calculated as a ring area weighted mean from recorded pith to bark wood density profiles.

![Study sites](graph_table/Study.sites.png){width=400px; height=400px}

Growth ring density and ring width measurements were taken from pith-to-bark, using an X-ray densitometer (Quintek Measurement Systems, TN) at a resolution of 0.0254 mm, and the data were reported as relative wood density on a dry weight basis. To verify the accuracy of dates and determine if there were missing rings, we used COFECHA, which is a dendrochronology software program used for cross-dating and quality control of annual ring measurements [@HOLMES1983].

@@

where RD is the mean ring density (kg.m^-3^), CA is cambial age (years), $b_0$, $b_1$ and $b_2$ are empirically determined parameters, and e is the residual error term. Parameter $b_0$ denotes the intercept when CA approaches a theoretical value of 0. We noticed that it was impossible to optimize the equation using the same estimated parameters for all tree cores, because the pattern of variation of annual rings average density of each tree is more or less different from the others. We, therefore, used the optim function of the stats package in the R [@R2018] statistical programming environment to find the best parameters iteratively for each tree core . The fit of the model was evaluated by visual analysis of the model residual curves against adjusted values and explanatory variables, fit indices (R^2^) [@Pape1999]. The graphic description in figure 2 of the three parameters of our model makes it possible to understand how each parameter influences our model (Fig 2).

@@

It can be observed in figure 2 that the parameters $b_0$ and $b_1$ influence the whole of the curve. An increase of one of these parameters leads to that of the density on all rings. Unlike parameter $b_0$ and $b_1$, parameter $b_2$ seems to be influenced by some of the density. A low value of this parameter causes a sharp increase of the density starting from the third ring and when this parameter increases a decrease of the density is observed at this level until the curve is similar to the cases of the parameters $b_0$ and $b_1$. Beyond the value 1, the parameter $b_2$ does not influence the variation of the density.\

## Determination of the wood density variability index
The variability index is an index to determine the variation in the density of pith-to-bark in order to identify tree families that are least sensitive to variations in climatic conditions. The residues of our model were used to obtain the wood density variability index, which was determined using an equation. We subtracted the density residues values of year $n-1$ from year $n$. We were interested more specifically in the difference between the average density residues of annual rings of two consecutive years, no matter whether the difference was positive or negative. The absolute value was used for each difference to get rid of the sign. The index was estimated from the following formula:

@@

Where V is the variability index, $D_n$ is the wood density residues of year $n$ and  $D_{n - 1}$ is the tree density residues of year $n-1$.

@@

The model we have developed has made it possible to optimize the density of the wood for each study site. In table 1, we observe that the average value of the parameters differs between Asselin and Saint-Casimir. for the parameters $b_0$ and $b_1$, the average is higher in Saint-Casimir and for parameter $b_2$ it is higher in Asselin.\

>>>>

---
title: "On the eletrodynamic of moving bodies"
author: "André Dogninema Soro, Jean-Romain Roussel, Alexis Achim"
output: 
  pdf_document: default
  word_document: default
bibliography: biblio.bibtex
editor_options: 
  chunk_output_type: inline
---

@@

Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of Canadian forest are harvested each year and an additional loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain resource constant, it is therefore essential to compensate the harvest and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better structural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \

@@

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood quality has a biologic and an economic component. Biologically, the resistance to mortality by vigour is important to ensure to harvest leaving trees. Economically the value of the wood for end-use is important for industries. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. More specifically the homogeneity of the density within the stem has an high importance. Its has been demonstrated that trees with a more uniform wood density within the stem are also the more resistant to mortality [@Hietz2013; @osazuwa2014] because a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. In fact, wood density uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bio energy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For example, a wide inter-annual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance [@x,@y]. It is therefore important from both a biological and an end-use points of view to have a wood with a uniform density within the stem. \

@@

The impact of climate is more perceptible on the properties of wood when there are wide variations of it such as drought, high temperatures or significant rainfall. The ability of trees to tolerate these changes in a tree species or within a species is often related to the interaction between genotype and environment. For example, studies have shown a low sensibility to genotype-related drought [@ferrell1966; @teskey1987]. Consequently, the sensitivity of trees to climate such as inter-annual variability in density can be controlled by genetics. It is therefore important to use trees that are less sensitive to climate variations because of their genome for breeding programs.

In this study we quantified and compared heterogeneity of wood density from trees of known parents in order to figure out if some individuals produce more heterogeneous wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the heterogeneity by comparing two growth sites we tried to characterize the environmental contribution to homogeneity. We ultimately aim to select and reproduce white spruce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogeneous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

@@

## Sampling
The dataset used in this study came from genetic selection for a series of tests according to provenances [@Beaulieu1996]. The improvement program for this region was led by Quebec’s Ministry of Forests, Wildlife and Parcs, Canada. Experimental plantings were initiated from white spruce seedlings grown in nurseries and planted at two years of age at a regular spacing between trees of 2 m in all directions. This series of tests was established in 1999 by the Canadian Forest Service and is one of the three essential parts of the white spruce breeding program in Quebec. Two sites were selected for sampling: one near Saint-Casimir, Quebec (46.6994, -72.1119), in the maple–basswood ecological zone (warm weather) , and the other one near Cabano (Asselin), Quebec (47.8808, -68.4436), in the balsam fir-yellow birch ecological zone (cool weather) (fig 1). Between 9 and 14 trees per site were sampled from each of the 93 full-sib families, for a total sample of 2365 trees. Tree height and diameter at breast height (DBH, 1.3 m above ground) were recorded in 2015 at the age of 15 years, and a wood increment core was extracted from the south facing side of each tree. Cores were stored in a freezer, conditioned to 7% moisture and cut to 1.68 mm thickness prior to X-ray densitometry analyses (Quintek Measurement Systems, TN) at a resolution of 0.0254 mm, and the data were reported as relative wood density on a dry weight basis. Wood density was calculated as a ring area weighted mean from recorded pith to bark wood density profiles.

![Map of Quebec showing the location of the two study sites where the data were collected](graph_table/Study.sites.png){width=300px; height=400px}

To verify the accuracy of dates and determine if there were missing rings, we used COFECHA, which is a dendrochronology software program used for cross-dating and quality control of annual ring measurements [@HOLMES1983]. thus, as the annual ring separation was carried out by an algorithm, we eliminated the samples (127 samples) whose dating was not correct and could not be modified.

@@

where RD is the mean ring density (kg.m^-3^), CA is cambial age (years), $b_0$, $b_1$ and $b_2$ are empirically determined parameters, and $e$ is the residual error term. Parameter $b_0$ denotes the intercept when CA approaches a theoretical value of 0. We noticed that it was impossible to optimize the equation using the same estimated parameters for all tree cores, because the pattern of variation of annual rings average density of each tree is more or less different from the others. We, therefore, used the optim function of the stats package in the R [@R2018] statistical programming environment to find the best parameters iteratively for each tree core . The fit of the model was evaluated by visual analysis of the model residual curves against adjusted values and explanatory variables, fit indices (R^2^) [@Pape1999]. The graphic description in figure 2 of the three parameters of our model makes it possible to understand how each parameter influences our model (Fig 2).

@@

It can be observed in figure 2 that the parameters $b_0$ and $b_1$ influence the whole of the curve. An increase of one of these parameters leads to that of the density on all the ring. Unlike parameter $b_0$ and $b_1$, parameter $b_2$ seems to be influenced by some of the density. A low value of this parameter causes a sharp increase of the density starting from the third growth year and when this parameter increases a decrease of the density is observed at this level until the curve is similar to the cases of the parameters $b_0$ and $b_1$. \

## Determination of the wood density variability index
The variability index is an index to determine the uniformity in the density of pith-to-bark in order to identify tree families that are least sensitive to variations in climatic conditions. The residues of our model were used to obtain the wood density variability index, which was determined using an equation. We subtracted the density residues values of year $n-1$ from year $n$. We were interested more specifically in the difference between the average density residues of annual rings of two consecutive years, no matter whether the difference was positive or negative. The absolute value was used for each difference to get rid of the sign. The index was estimated from the following formula:

@@

Where V is the variability index, $D_n$ is the wood density residues of year $n$ and  $D_{n - 1}$ is the wood density residues of year $n-1$.

@@

The model we have developed has made it possible to optimize the density of the wood for each study site. In table 1, we observe that the average value of the parameters differs between Asselin and Saint-Casimir. For the parameters $b_0$ and $b_1$, the average is higher in Saint-Casimir and for parameter $b_2$ it is higher in Asselin.\
